---
title: "RR Week 2 assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set working directory and read/transform steps database
```{r, echo = TRUE}
setwd("~/Documents/Documenten/R/Reproducible Research/Assignment week 2")
activity <- read.csv("activity.csv")
activity <- transform(activity, date = as.Date(activity$date))
activity <- transform(activity, steps = as.numeric(activity$steps))
activity <- transform(activity, interval = as.numeric(activity$interval))
```

##Histogram of steps total
```{r, echo = TRUE}
hist(activity$steps, xlab = "number of steps", ylab = "amount of steps", col = "green")
```

##Mean and median number of steps
```{r, echo = TRUE}
mean_steps <- tapply(activity$steps, activity$date, mean)
median_steps <- tapply(activity$steps, activity$date, median)
frame <- data.frame(mean_steps, median_steps)
frame
```

##Plot of steps on average each day
```{r, echo = TRUE}
agg <- aggregate(steps ~ date, activity, mean)
plot(agg$date, agg$steps, type = "l")
```

##Which interval has most steps on average per day?
```{r, echo = TRUE}
agg_int <- aggregate(.~ interval, activity, max)
agg_int[which.max(agg_int$steps), ]
```

##Missing values and imputation
```{r, echo = TRUE}
table(is.na(activity))
library(Hmisc)
activity_noNA <- activity
activity_noNA$steps <- with(activity_noNA, impute(steps, mean))
head(activity_noNA)
```

##Histogram for imputed set and check if there are differences in mean and median values afterwards
```{r, echo = TRUE}
hist(activity_noNA$steps, xlab = "number of steps", ylab = "frequency", col = "green")
```

We see that the there is a higher frequency of a low number of steps per interval.

```{r, echo = TRUE}
mean_steps_noNA <- tapply(activity_noNA$steps, activity_noNA$date, mean)
median_steps_noNa <- tapply(activity_noNA$steps, activity_noNA$date, median)
frame_noNA <- data.frame(mean_steps_noNA, median_steps_noNa)
head(frame_noNA)
diff <- cbind(frame, frame_noNA)
head(diff)
```

We also see that there is no difference between the mean and median values for both sets.

##Difference in amount of steps per interval for weekend or weekdays?
```{r, echo = TRUE}
weekday <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
activity_noNA$daytype <- factor((weekdays(activity_noNA$date) %in% weekday), levels = c(FALSE, TRUE), labels = c("Weekend", "Weekday"))
agg_day <- aggregate(steps ~ interval + daytype, activity_noNA, mean)
```

Save the plot to a file!

```{r, echo = TRUE}
g <- ggplot(data = agg_day, aes(interval, steps))
p <- g + geom_line(color = "red") + facet_grid(daytype~.) + labs(title = "Average steps per interval on weekdays or in the weekend")
p;
ggsave('steps.png', width = 16, height = 9, dpi = 100)
```
